---
title: "Stat 245 -- Binary Regression Practice"
author: "A. Student"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    fig_height: 2.2
    fig_width: 4
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```


```{r}
searches <- read_csv('https://sldr.netlify.app/data/election_searches.csv') |>
  mutate(Race_Ethnicity = str_remove_all(Race_Ethnicity, '_NA'),
         Race_Ethnicity = str_remove_all(Race_Ethnicity, 'NA_')) |>
  mutate(Race_Ethnicity = ifelse(grepl(pattern = '_', Race_Ethnicity), 
                                 'Multiple', 
                                 Race_Ethnicity)) |>
  filter(Education != 'Less than 9th grade' & Sex != 'Other')   |> # remove as there are only very few
  mutate(Income = fct_relevel(Income,
                              "Under $10,000",
                              "$10,000 to $14,999",
                              "$15,000 to $19,999",
                              "$20,000 to $29,999",
                              "$30,000 to $39,999",
                              "$40,000 to $49,999",
                              "$50,000 to $74,999",
                              "$75,000 to $99,999",
                              "$100,000 to $149,999",
                              "$150,000 and over",
                              "I prefer not to say"),
         Education = fct_relevel(Education,
                                 "9th to 12th grade",
                                 "High school graduate",
                                 "Some college or associate degree",
                                 "Bachelors", 
                                 "Advanced" ),
         Post_Choice = factor(Post_Choice),
         Post_Choice = fct_relevel(Post_Choice, "Tony Abbott", "Julia Gillard"))  

glimpse(searches)
```

### Model Planning 

Response: Post_Choice 
Predictors: Searches, Pre_Choice, Trust_Search 
Interactions: Group * Searches, Group * Pre-Choice, Group * Trust_Search, Group * Party 

### Data Exploration 

### Model Fitting 

```{r}
search.logr <- glm(factor(Post_Choice) ~ Group * Searches + Pre_Choice + Trust_Search + Party, data = searches, family = binomial(link='logit'))

summary(search.logr)
```

### Model Assessment 

Linearity?

## Independence of Residuals 

```{r}
s245::gf_acf(resids(~search.logr))
```


## Mean-Variance 

```{r}
require(DHARMa)
search_sim <- simulateResiduals(search.logr)

gf_point(search_sim$scaledResiduals ~ rank(predict(search.logr)),
         alpha = 0.2) |>
  gf_labs(x = 'Fitted Values',
          y = 'Scaled Residuals')
```
The residuals are uniformly distributed which is why we can say that the mean-variance relationship condition passes in this case

### Model Selection 

```{r}
car::Anova(search.logr)
```

### Interpretation and Prediction Plots 

```{r}
require(ggeffects)
pred_plot_data <- ggpredict(search.logr, 
          terms = 'Party')
plot(pred_plot_data)
```






  