---
title: "Diversity in PhD Theses"
author: "Sharon Velpula"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
classoption: landscape
header-includes:
- \usepackage[fontsize=20pt]{scrextend}
---

```{r setup, include=FALSE}
require(tidyverse)
require(mosaic)
require(ggformula)
knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      fig.width = 7, 
                      fig.height = 4)

theme_set(theme_minimal(base_size = 16))
```


<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## Data

```{r}
pi <- read_csv('https://sldr.netlify.app/data/phd_innovation.csv')

pi
```

```{r}
gf_histogram(~ uptake, fill = ~ race_ethnicity, bins = 15, data = pi)
```

I chose a histogram because I wanted to visualize both the categorical variable race_ethnicity and the quantitative variable uptake. We can see that up till a certain point, the increase in uptake shows an increase in the count of different ethnicity groups. However, after that point, increasing uptake shows a decrease in the count of the ethnicity groups. Just by observation, it looks like most people experience an uptake between 1 and 1.5, with the Asians experiencing it the most. 





  