---
title: "Stat 245 -- Michigan Housing"
author: "Sharon Velpula"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for homework, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```

## Read in Data
```{r}
hdma_mi_20 <- read_csv('https://sldr.netlify.app/data/hdma-mi-20.csv', show_col_types = FALSE)

head(hdma_mi_20)
```

## Data Wrangling
```{r}
hdma_mi_20 <- hdma_mi_20 |>
  mutate(action_taken = case_when(action_taken == 1 ~ 'Loan granted', action_taken == 3 ~ 'Loan denied' ))

mosaic::tally(~action_taken, data = hdma_mi_20)
```

```{r}
hdma_mi_20_sex <- hdma_mi_20 %>% 
  filter(derived_sex %in% c("Male", "Female", "Joint", "Sex Not Available"))
```


## Calculate average denial rate for every sex
```{r}
hdma_mi_20_denied <- hdma_mi_20 |> 
  group_by(derived_sex) |> 
  mutate(rate_of_denial=paste0(round(count(action_taken=='Loan denied')/sum(action_taken=='Loan granted', action_taken=='Loan denied')*100))) %>%

select("derived_sex", "rate_of_denial")

glimpse(hdma_mi_20_denied)
```

```{r}
unique(hdma_mi_20_denied$rate_of_denial)
```

## New Dataframe including only sex and denied numbers
```{r}
data <- data.frame(
  sex = c("Male", "Female", "Joint", "Sex Not Available"), 
  denied = c(8, 5, 9, 13)
)
```

## Graph
```{r}
data %>%
  ggplot() +
  aes(y= sex, x = denied) + 
  geom_col(position = "dodge") +
  theme_minimal()
```

I attempted to recreate the line graph that shows the rate of denial in response to credit score by The Markup. But while going through the process, I wanted to show the rate of denial in response to credit score by sex. But I was not able to do all the calculations properly so my end result is basically a bar plot showing denied numbers by sex. Looks like those whose sex was not available were denied the most while females were denied the least. 

  