---
title: "Stat 245 -- RRStudio-Jump Start"
author: "Sharon Velpula"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
classoption: landscape
header-includes:
- \usepackage[fontsize=20pt]{scrextend}
---

```{r setup, include=FALSE}
require(tidyverse)
require(mosaic)
require(ggformula)
knitr::opts_chunk$set(echo = TRUE,
                      error = TRUE,
                      fig.width = 7, 
                      fig.height = 4)

theme_set(theme_minimal(base_size = 16))
```


<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## Read in Data 

```{r}
wrdata <- 
  readr::read_csv('https://sldr.netlify.app/data/sustainable-livelihoods.csv', 
                  show_col_types = FALSE)

gf_percents(~ Leadership_Program, fill = ~ Management_Confidence, 
            data = wrdata, na.rm = TRUE)
```

### Question

How did participation in the Leadership program affect the management confidence?

I chose a bar graph because Management_Confidence and Leadership_Program are categorical variables. I wanted to visualize the difference in the percentage of people's participation in the Leadership progam. Since there is a visible difference in the percentage, a bar graph was my first choice. Furthermore, I made it a stacked bar graph to show the difference in Management_Confidence within those who participated in the leadership program and those who didn't. A stacked bar graph makes it easier to estimate the percentage within a category. 

We can see that those who participated in the leadership program had more management confidence than those who didn't participate. 


I think I could use a pie graph to visualize these variables. The percentage of those who participated in the leadership program and those who did not, adds up to 100% and the comparison will be more evident. 
  