---
title: "Stat 243 -- another R Markdown doc"
author: "Sharon Velpula"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    fig_height: 2.2
    fig_width: 4
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
require(mosaic)      # this loads ggformula (for plotting), etc. too
require(Lock5withR)  # this loads data sets from our book

# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```




1. Scatterplot with best fit overlain 
```{r}
gf_point(Price ~ PPM, data = InkjetPrinters) %>% gf_lm()
```



2. Equation of best fit line 
```{r}
lm(Price ~ PPM, data = InkjetPrinters)
```
$$ \widehat{Price} = 90.88\times\mbox{PPM} - 94.22 $$



3. Diagnostic plots
```{r}
lmResult <- lm(Price ~ PPM, data = InkjetPrinters)
plot(lm(Price ~ PPM, data = InkjetPrinters))
```


For the first diagnostic plot, we can see that the residuals are almost equally spread out which means that there aren't any non-linear relationships. 

For the second diagnostic plot, we can see that the residuals almost follow a straight line and appear to deviate at some points consistently. So, the residuals are normally distributed. 

Therefore, SLM holds for this model. 



4. Model Utility Test
$$ H_0: \beta_0 = 0 $$
$$ H_a: \beta_1\ne 0 $$

```{r}
summary(lm(Price ~ PPM, data = InkjetPrinters))
```

From what we can see, the test statistic is 4.663 and the p value is 0.0001934. 
Since the pvalue is very small, we can reject the null hypothesis and say that 
we have convincing evidence that PPM is an effective predictor of price in ink
jet printers. 



5. Coefficient of Determination
```{r}
summary(lm(Price ~ PPM, data = InkjetPrinters))
```
PPM explains 54.7% variability in Price of these 20 printers



6. Confidence Interval 
```{r}
tstar <- qt(0.975, df = 18)
lm(Price ~ PPM, data = InkjetPrinters)$coefficients[2]+c(-1, 1)*tstar*19.49
```
Since we rejected the null hypothesis, as expected, our pvalue of 0.0001934 is less than 0.05.
Therefore the true slope lies between 49.9311 and 131.8250 and the slope given by our model 
also lies within this interval. 
